---
title: "House Price Prediction with R Part I"
author: "Md Mominul Islam"
date: "11/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## House Price Prediction

Prediction in research satisfies one of humankind's greatest primal desires. We make a lot of predictions to see into the future and know what awaits us. Most scientists regularly use prediction in research as a fundamental of the scientific method, when they generate a hypothesis and predict what will happen. Machine learning models function similarly to humans. They use instances/logics they've seen before to forecast what will happen given a certain set of circumstances.

For this lesson, we will bulid a model to figure out how much a house will sell for?

The steps that we will follow are given below:

1. Load Package

2. Read Train and Test data

3. Combine Train and Test dataset

4. Data Visualization

5. Missing Value

6. Removing Skewed Variables

7. Build the model

8. Variable importance

9. Final Prediction

10. Calculate RMSE

## Loading and Exploring Data

## Loading libraries required and reading the data into R

Loading R packages used besides base R.

```{r, echo = TRUE,message=FALSE, warning=FALSE}
library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
library(xgboost)
```

The first thing we want to do is familiarize ourselves with the data. By checking the dataset, we can have a holistic picture about the variables that are present in the dataset.

```{r}
train.data <- read.csv(("Data Set/train.csv"), stringsAsFactors = F)
test.data <- read.csv(("Data Set/test.csv"), stringsAsFactors = F)
```

## Structure of the data

```{r,results="hide"}
dim(train.data)
str(train.data)
dim(test.data)
str(test.data)
```

The housing train data set has 1460 rows and 81 features with the target feature "SalePrice". The housing test data set has 1459 rows and 80 features with the target feature Sale Price.

```{r, message=FALSE, warning=FALSE}
#Count the number of columns that consists of text data
sum(sapply(train.data[,1:81], typeof) == "character")
#Count the number of columns that consists of numerical data
sum(sapply(train.data[,1:81], typeof) == "integer")
```

We have 43 columns that consist of character variable and 38 columns are numerical. The text data could be challenging to work with. For those that are numerical,let's take a look at some descriptive statistics.

```{r, message=FALSE, warning=FALSE}
cat('Train data has', dim(train.data)[1], 'rows and', dim(train.data)[2], 'columns.')
cat('Test data has', dim(test.data)[1], 'rows and', dim(test.data)[2], ' columns.')

# The percentage of data missing in train
tr.miss <- (sum(is.na(train.data)) / (nrow(train.data) *ncol(train.data)))*100
cat("Missing value Percentage for Train Data is", tr.miss, '%')
# The percentage of data missing in test
tst.miss <- (sum(is.na(test.data)) / (nrow(test.data) * ncol(test.data)))*100
cat("Missing value Percentage for Train Data is", tst.miss, '%')
# Check for duplicated rows
cat("The number of duplicated rows are", nrow(train.data) - nrow(unique(train.data)))
```

We have 5.88% missing values in our train data set and 7.15% missing values in our test data set.

## Combine data

Since test data set has no "Saleprice" variable. We will create it and then combine.

```{r, message=FALSE, warning=FALSE}
test.data$SalePrice<-rep(NA,1459)
combined_data <-bind_rows(train.data,test.data)
all <- rbind(train.data,test.data)
dim(all)
```

After combining the data set, we can clearly see that there are 81 columns in our data set.

## Data Exploration

### The response variable: SalePrice

```{r, message=FALSE}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="blue", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

As we can see, the sale prices are right skewed. This was expected as few people can afford very expensive houses. We will keep this in mind, and take measures before modeling.

### Summary Statistics for SalePrice

```{r, message=FALSE, warning=FALSE}
kable(as.array(summary(all$SalePrice)),
      caption = 'Summary Statistics of Sales Price')
```


## Important Numeric Predictors

The character variables need some work before we can use them. To get a feel for the dataset, I decided to first see which numeric variables have a high correlation with the SalePrice.

### Correlations with SalePrice

```{r, message=FALSE, warning=FALSE}
numericVars <- which(sapply(all, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
all_numVar <- all[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs") #correlations of all numeric variables
#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]
corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")
```

We have 38 numerical variables in our data set.Altogether, there are 10 numeric variables with a correlation of at least 0.5 with SalePrice. All those correlations are positive. As from the correlation matrix plot, we can see that  OverallQual, GrLivArea, GarageCars are highly correlated with the SalePrice. We have to keep in mind about multicollinearity. For example: the correlation between GarageCars and GarageArea is very high (0.89), and both have similar (high) correlations with SalePrice. 

The other 6 six variables with a correlation higher than 0.5 with SalePrice are:

-TotalBsmtSF: Total square feet of basement area

-1stFlrSF: First Floor square feet

-FullBath: Full bathrooms above grade

-TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)

-YearBuilt: Original construction date

-YearRemodAdd: Remodel date (same as construction date if no remodeling or additions)

### Overall Quality

Overall Quality has the highest correlation with SalePrice among the numeric variables (0.79). It rates the overall material and finish of the house on a scale from 1 (very poor) to 10 (very excellent).

```{r, message=FALSE, warning=FALSE}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot(col='red') + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

The positive correlation is certainly there indeed, and seems to be a slightly upward curve. Regarding outliers, we do not see any extreme values. If there is a candidate to take out as an outlier later on, it seems to be the expensive house with grade 4.

### Above Grade (Ground) Living Area (square feet)

The numeric variable with the second highest correlation with SalesPrice is the Above Grade Living Area. This make a lot of sense; big houses are generally more expensive.

```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='red') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_text_repel(aes(label = ifelse(all$GrLivArea[!is.na(all$SalePrice)]>4500, rownames(all), '')))
        
```

Especially the two houses with really big living areas and low SalePrices seem outliers (houses 524 and 1299, see labels in graph). we will not take them out yet, as taking outliers can be dangerous. For instance, a low score on the Overall Quality could explain a low price. However, as we can see below, these two houses actually also score maximum points on Overall Quality. Therefore, we will keep houses 1299 and 524 in mind as prime candidates to take out as outliers.

```{r}
#Looking at highly correlated values of outliers
a.outlr <- all[c(524, 1299), c('SalePrice', 'GrLivArea', 'OverallQual')]
kable(a.outlr)
```


### Garage Cars

```{r, message=FALSE, warning=FALSE}

ggplot(all, aes(x = GarageCars, y = SalePrice)) +
  geom_point(alpha = .5) +
  geom_smooth()+
  scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
  
```

As we can see, if there in no garage for cars then the price is low. With the increase of GarageCars, SalePrice increases.

# Missing data, label encoding, and factorizing variables

## Completeness of the data

First of all, we would like to see which variables contain missing values.

```{r}
NAcol <- which(colSums(is.na(all)) > 0)
sort(colSums(sapply(all[NAcol], is.na)), decreasing = TRUE)
cat('There are', length(NAcol), 'columns with missing values')
```

We can clearly see that there are 1459 NAs in SalePrice which is matched perfectly with the size of the test set. So, we have to fix NAs in 34 predictor variables.

## Missing Value Imputation

We are going to fix the 34 predictors that have missing values in this area. We will go over them one by one, starting with the most common NAs and working our way down until we've repaired them all. If we come across a variable that is part of a group with other variables, we shall treat them as a unit. Pool, Garage, and Basement, for example, each have a number of variations.

We turned character variables into ordinal integers if there is evident ordinality, or into factors if levels are categories without ordinality, in addition to ensuring that NAs are handled. We'll use one-hot encoding to transform these factors to numbers later (using the model.matrix function).

**Pool Quality and the PoolArea variable**
  
The PoolQC is the variable with most NAs. 


```{r}
kable(unique(all$PoolQC),
      caption= 'Pool Quality Features')
```

The description is as follows:
  
PoolQC: Pool quality,Ex:Excellent, Gd:Good,TA: Average/Typical,Fa:Fair, NA:No Pool

So, it is obvious that we need to just assign 'No Pool' to the NAs. Also, the high number of NAs makes sense as normally only a small proportion of houses have a pool.

```{r}
all$PoolQC[is.na(all$PoolQC)] <- 'None'
```

It is also clear that we can label encode this variable as the values are ordinal. As there a multiple variables that use the same quality levels, we are going to create a vector that we can reuse later on.

```{r}
Qualities <- c('None' = 0, 'Fa' = 1, 'Gd' = 2, 'Ex' = 3)
```

Now, we can use the function 'revalue' to do the work for me.

```{r, message=FALSE}

all$PoolQC<-as.integer(revalue(all$PoolQC, Qualities))
table(all$PoolQC)
```

However, there is a second variable that relates to Pools. This is the PoolArea variable (in square feet). As you can see below, there are 3 houses without PoolQC. First, I checked if there was a clear relation between the PoolArea and the PoolQC. As I did not see a clear relation (bigger of smaller pools with better PoolQC), I am going to impute PoolQC values based on the Overall Quality of the houses (which is not very high for those 3 houses).

```{r}
ql <- all[all$PoolArea>0 & all$PoolQC==0, c('PoolArea', 'PoolQC', 'OverallQual')]
kable(ql)
```
### Miscellaneous Feature
  
Within Miscellaneous Feature, there are 2814 NAs. As the values are not ordinal, we will convert MiscFeature into a factor. Values:	

```{r}
unique(all$MiscFeature)
```


Elev	Elevator

Gar2	2nd Garage (if not described in garage section)

Othr	Other

Shed	Shed (over 100 SF)

TenC	Tennis Court

NA	  None

```{r,warning=FALSE,message=FALSE}
all$MiscFeature[is.na(all$MiscFeature)] <- 'None'
all$MiscFeature <- as.factor(all$MiscFeature)

ggplot(all[!is.na(all$SalePrice),], aes(x=MiscFeature, y=SalePrice)) +
  geom_bar(stat='summary', fun.y = "median", fill='blue') +
  scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..))

kable(table(all$MiscFeature),
      caption= 'Table for Miscelenious Features',
      col.names = c('Variable','Frequency'))
```

When looking at the frequencies, the variable seems irrelevant to me. Having a shed probably means 'no Garage', which would explain the lower sales price for Shed. Also, while it makes a lot of sense that a house with a Tennis court is expensive, there is only one house with a tennis court in the training set.

### Alley Access Type to property

```{r}
unique(all$Alley)
```
  

  
Within Alley, there are 2721 NAs. As the values are not ordinal, we will convert Alley into a factor. 

Grvl	Gravel

Pave	Paved

NA 	No alley access

```{r,warning=FALSE,message=FALSE}
all$Alley[is.na(all$Alley)] <- 'None'
all$Alley <- as.factor(all$Alley)

ggplot(all[!is.na(all$SalePrice),], aes(x=Alley, y=SalePrice)) +
  geom_bar(stat='summary', fun.y = "median", fill='blue')+
  scale_y_continuous(breaks= seq(0, 200000, by=50000), labels = comma)
kable(table(all$Alley),
      caption= 'Table for Alley variable',
      col.names = c('Variable','Frequency'))
```

### Fence Quality
  
We have seen that, there are 2348 NAs within Fence variable. 

```{r}
unique(all$Fence)
```
  
*Meaning*

GdPrv	Good Privacy

MnPrv	Minimum Privacy

GdWo	Good Wood

MnWw	Minimum Wood/Wire

NA	No Fence

```{r}
all$Fence[is.na(all$Fence)] <- 'None'
kable(table(all$Fence),
      caption= 'Table for Fence Quality',
      col.names = c('Type','Frequency'))
# Summary Based on median
all[!is.na(all$SalePrice),] %>% 
  group_by(Fence) %>% 
  summarise(median = median(SalePrice), counts=n())
```

My conclusion is that the values do not seem ordinal (no fence is best). Therefore, I will convert Fence into a factor.

```{r}
all$Fence <- as.factor(all$Fence)
```

### Fireplace variables

**Fireplace quality**

```{r, message=F, warning=F}
all$FireplaceQu[is.na(all$FireplaceQu)] <- 'None'
all$FireplaceQu<-as.integer(revalue(all$FireplaceQu, Qualities))
table(all$FireplaceQu)
```
  
Values:
  
Ex	Excellent - Exceptional Masonry Fireplace

Gd	Good - Masonry Fireplace in main level

TA	Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement

Fa	Fair - Prefabricated Fireplace in basement

Po	Poor - Ben Franklin Stove

NA	No Fireplace

**Number of fireplaces**

Fireplaces is an integer variable, and there are no missing values.

```{r}
table(all$Fireplaces)
sum(table(all$Fireplaces))
```

###Lot variables
  
3 variables. One with 1 NA, and 2 complete variables.

**LotFrontage: Linear feet of street connected to property**
  
486 NAs. The most reasonable imputation seems to take the median per neigborhood.

```{r}
ggplot(all[!is.na(all$LotFrontage),], aes(x=as.factor(Neighborhood), y=LotFrontage)) +
  geom_bar(stat='summary', fun.y = "median", fill='blue') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**LotShape: General shape of property**
  
No NAs. Values seem ordinal (Regular=best)

Reg	Regular

IR1	Slightly irregular

IR2	Moderately Irregular

IR3	Irregular

```{r}
all$LotShape<-as.integer(revalue(all$LotShape, c('IR3'=0, 'IR2'=1, 'IR1'=2, 'Reg'=3)))
table(all$LotShape)
sum(table(all$LotShape))
```

**LotConfig: Lot configuration**
  
No NAs. The values seemed possibly ordinal to me, but the visualization does not show this. Therefore, I will convert the variable into a factor.

Inside	Inside lot

Corner	Corner lot

CulDSac	Cul-de-sac

FR2	Frontage on 2 sides of property

FR3	Frontage on 3 sides of property

```{r}
ggplot(all[!is.na(all$SalePrice),], aes(x=as.factor(LotConfig), y=SalePrice)) +
  geom_bar(stat='summary', fun.y = "median", fill='blue')+
  scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
  geom_label(stat = "count", aes(label = ..count.., y = ..count..))
```

```{r}
all$LotConfig <- as.factor(all$LotConfig)
table(all$LotConfig)
sum(table(all$LotConfig))
```

### Garage variables




GarageType
GarageYrBlt
GarageFinish
GarageCars
GarageArea
GarageQual
GarageCond

**GarageType: Garage location**
  
The values do not seem ordinal, so I will convert into a factor.

2Types	More than one type of garage
Attchd	Attached to home
Basment	Basement Garage
BuiltIn	Built-In (Garage part of house - typically has room above garage)
CarPort	Car Port
Detchd	Detached from home
NA	No Garage

```{r}
all$GarageType[is.na(all$GarageType)] <- 'No Garage'
all$GarageType <- as.factor(all$GarageType)
table(all$GarageType)
```

**GarageFinish: Interior finish of the garage**
  
The values are ordinal.

Fin	Finished

RFn	Rough Finished

Unf	Unfinished

NA	No Garage       

```{r}
all$GarageFinish[is.na(all$GarageFinish)] <- 'None'
Finish <- c('None'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)

all$GarageFinish<-as.integer(revalue(all$GarageFinish, Finish))
table(all$GarageFinish)
```

As from the data we can see that there are 159 houses with no garage, 1230 houses with unfinished garage.

**GarageQual: Garage quality**
  
Another variable than can be made ordinal with the Qualities vector.

Ex	Excellent

Gd	Good

TA	Typical/Average

Fa	Fair

Po	Poor

NA	No Garage

```{r}
all$GarageQual[is.na(all$GarageQual)] <- 'None'
Ga.Quality <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
all$GarageQual<-as.integer(revalue(all$GarageQual, Ga.Quality))
table(all$GarageQual)
```

**GarageCond: Garage condition**
  
Garage Condition is another categorical variable which can be converted to integer.

```{r}
all$GarageCond[is.na(all$GarageCond)] <- 'None'
all$GarageCond<-as.integer(revalue(all$GarageCond, Qualities))
table(all$GarageCond)
```


### MSSubClass
         
MSSubClass: Identifies the type of dwelling involved in the sale.	

20	1-STORY 1946 & NEWER ALL STYLES

30	1-STORY 1945 & OLDER

40	1-STORY W/FINISHED ATTIC ALL AGES

45	1-1/2 STORY - UNFINISHED ALL AGES

50	1-1/2 STORY FINISHED ALL AGES

60	2-STORY 1946 & NEWER

70	2-STORY 1945 & OLDER

75	2-1/2 STORY ALL AGES

80	SPLIT OR MULTI-LEVEL

85	SPLIT FOYER

90	DUPLEX - ALL STYLES AND AGES

120	1-STORY PUD (Planned Unit Development) - 1946 & NEWER

150	1-1/2 STORY PUD - ALL AGES

160	2-STORY PUD - 1946 & NEWER

180	PUD - MULTILEVEL - INCL SPLIT LEV/FOYER

190	2 FAMILY CONVERSION - ALL STYLES AND AGES

These classes are coded as numbers, but really are categories.
         
```{r}
str(all$MSSubClass)
all$MSSubClass <- as.factor(all$MSSubClass)
#revalue for better readability
all$MSSubClass<-revalue(all$MSSubClass, c('20'='1 story 1946+', '30'='1 story 1945-', '40'='1 story unf attic', '45'='1,5 story unf', '50'='1,5 story fin', '60'='2 story 1946+', '70'='2 story 1945-', '75'='2,5 story all ages', '80'='split/multi level', '85'='split foyer', '90'='duplex all style/age', '120'='1 story PUD 1946+', '150'='1,5 story PUD all', '160'='2 story PUD 1946+', '180'='PUD multilevel', '190'='2 family conversion'))
kable(table(all$MSSubClass))
```
         
### Above Ground Living Area, and other surface related variables (in square feet)
         
As we have already visualized the relation between the Above Ground Living Area and SalePrice in my initial explorations, we will now just display the distribution itself. As there are more 'square feet' surface measurements in the Top 20, we are taking the opportunity to bundle them in this section. Note: GarageArea is taken care of in the Garage variables section.We are also adding 'Total Rooms Above Ground' (TotRmsAbvGrd) as this variable is highly correlated with the Above Ground Living Area(0.81).

```{r, warning=FALSE, message=FALSE, out.width="100%"}
sca1 <- ggplot(data= all, aes(x=GrLivArea)) +
 geom_density() + labs(x='Square feet living area')
sca2 <- ggplot(data=all, aes(x=as.factor(TotRmsAbvGrd))) +
 geom_histogram(stat='count') + labs(x='Rooms above Ground')
sca3 <- ggplot(data= all, aes(x=X1stFlrSF)) +
 geom_density() + labs(x='Square feet first floor')
sca4 <- ggplot(data= all, aes(x=X2ndFlrSF)) +
 geom_density() + labs(x='Square feet second floor')
sca5 <- ggplot(data= all, aes(x=TotalBsmtSF)) +
 geom_density() + labs(x='Square feet basement')
sca6 <- ggplot(data= all[all$LotArea<100000,], aes(x=LotArea)) +
 geom_density() + labs(x='Square feet lot')
sca7 <- ggplot(data= all, aes(x=LotFrontage)) +
 geom_density() + labs(x='Linear feet lot frontage')
sca8 <- ggplot(data= all, aes(x=LowQualFinSF)) +
 geom_histogram() + labs(x='Low quality square feet 1st & 2nd')

layout <- matrix(c(1,2,5,3,4,8,6,7),4,2,byrow=TRUE)
multiplot(sca1, sca2, sca3, sca4, sca5, sca6, sca7, sca8, layout=layout)
```

###The most important categorical variable; Neighborhood
         
Th first graph shows the median SalePrice by Neighorhood. The frequency (number of houses) of each Neighborhood in the train set is shown in the labels.

The second graph below shows the frequencies across all data.

```{r, warning=FALSE, out.width="100%"}
nei1 <- ggplot(all[!is.na(all$SalePrice),], aes(x=Neighborhood, y=SalePrice)) +
 geom_bar(stat='summary', fun.y = "median", fill='blue') +
 theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
 scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
 geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
 geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
nei2 <- ggplot(data=all, aes(x=Neighborhood)) +
 geom_histogram(stat='count')+
 geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3)+
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(nei1, nei2)
```

### Overall Quality, and other Quality variables
         
I have already visualized the relation between Overall Quality and SalePrice in my initial explorations, but I want to visualize the frequency distribution as well. As there are more quality measurements, I am taking the opportunity to bundle them in this section.

```{r, warning=FALSE, message=FALSE, out.width="100%"}
q1 <- ggplot(data=all, aes(x=as.factor(OverallQual))) +
 geom_histogram(stat='count')
q2 <- ggplot(data=all, aes(x=as.factor(ExterQual))) +
 geom_histogram(stat='count')
q3 <- ggplot(data=all, aes(x=as.factor(BsmtQual))) +
 geom_histogram(stat='count')
q4 <- ggplot(data=all, aes(x=as.factor(KitchenQual))) +
 geom_histogram(stat='count')
q5 <- ggplot(data=all, aes(x=as.factor(GarageQual))) +
 geom_histogram(stat='count')
q6 <- ggplot(data=all, aes(x=as.factor(FireplaceQu))) +
 geom_histogram(stat='count')
q7 <- ggplot(data=all, aes(x=as.factor(PoolQC))) +
 geom_histogram(stat='count')

layout <- matrix(c(1,2,8,3,4,8,5,6,7),3,3,byrow=TRUE)
multiplot(q1, q2, q3, q4, q5, q6, q7, layout=layout)

```

Overall Quality is crucial, yet it is also more detailed than the other characteristics. External quality is important as well, but it has a strong link to overall quality (0.73). Kitchen quality appears to be something to bear in mind, given all residences have a kitchen and there is considerable variation in terms of content. The majority of garages have Q3 quality, which does not appear to make much of a difference. Fireplace Quality is on the list of essential variables and has a good association. The PoolQC is simply lacking (the 13 pools cannot even be seen on this scale). 
         
### The second most important categorical variable; MSSubClass

The first visualization shows the median SalePrice by MSSubClass. The frequency (number of houses) of each MSSubClass in the train set is shown in the labels.

The histrogram shows the frequencies across all data. Most houses are relatively new, and have one or two stories.

```{r, warning=FALSE, out.width="100%"}
ms1 <- ggplot(all[!is.na(all$SalePrice),], aes(x=MSSubClass, y=SalePrice)) +
 geom_bar(stat='summary', fun.y = "median", fill='blue') +
 theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
 scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
 geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
 geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
ms2 <- ggplot(data=all, aes(x=MSSubClass)) +
 geom_histogram(stat='count')+
 geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(ms1, ms2)
```

###Garage variables
         
Several Garage variables have a high correlation with SalePrice, and are also in the top-20 list of the quick random forest. However, there is multicolinearity among them and I think that 7 garage variables is too many anyway.We feel that something like 3 variables should be sufficient (possibly GarageCars, GarageType, and a Quality measurement), but before I do any selection I am visualizing all of them in this section.

```{r}
#correct error
all$GarageYrBlt[2593] <- 2007 #this must have been a typo. GarageYrBlt=2207, YearBuilt=2006, YearRemodAdd=2007.
```


```{r, warning=FALSE, message=FALSE, out.width="100%"}
g1 <- ggplot(data=all[all$GarageCars !=0,], aes(x=GarageYrBlt)) +
 geom_histogram()
g2 <- ggplot(data=all, aes(x=as.factor(GarageCars))) +
 geom_histogram(stat='count')
g3 <- ggplot(data= all, aes(x=GarageArea)) +
 geom_density()
g4 <- ggplot(data=all, aes(x=as.factor(GarageCond))) +
 geom_histogram(stat='count')
g5 <- ggplot(data=all, aes(x=GarageType)) +
 geom_histogram(stat='count')
g6 <- ggplot(data=all, aes(x=as.factor(GarageQual))) +
 geom_histogram(stat='count')
g7 <- ggplot(data=all, aes(x=as.factor(GarageFinish))) +
 geom_histogram(stat='count')

layout <- matrix(c(1,5,5,2,3,8,6,4,7),3,3,byrow=TRUE)
multiplot(g1, g2, g3, g4, g5, g6, g7, layout=layout)

```

GarageCars and GarageArea are highly correlated. Here, GarageQual and GarageCond also seem highly correlated, and both are dominated by level =3. 

# Feature engineering

## Total number of Bathrooms

There are 4 bathroom variables. Individually, these variables are not very important. However, I assume that I if I add them up into one predictor, this predictor is likely to become a strong one.

"A half-bath, also known as a powder room or guest bath, has only two of the four main bathroom components-typically a toilet and sink." Consequently, I will also count the half bathrooms as half.

```{r}
all$TotBathrooms <- all$FullBath + (all$HalfBath*0.5) + all$BsmtFullBath + (all$BsmtHalfBath*0.5)
```

As you can see in the first graph, there now seems to be a clear correlation (it's 0.63). The frequency distribution of Bathrooms in all data is shown in the second graph.

```{r, warning=FALSE}
tb1 <- ggplot(data=all[!is.na(all$SalePrice),], aes(x=as.factor(TotBathrooms), y=SalePrice))+
  geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
  scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
tb2 <- ggplot(data=all, aes(x=as.factor(TotBathrooms))) +
  geom_histogram(stat='count')
grid.arrange(tb1, tb2)
```
         


